Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-22 aistats.tex
Add figure for MAP upper bound

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
Let us recall the form of the divergence with different parametrizations.
For a Gaussian with known mean $0$ and covariance $\sigma^2$, 
natural and mean parameters $\theta, \m$;
\begin{align}
	\KL( \sigma_*^2 ; \sigma_n^2 ) 
	&= \half \left ( \frac{\sigma_*^2}{ \sigma_n^2} - 1 - \log \frac{\sigma_*^2}{ \sigma_n^2} \right)
	= \phi\paren{\frac{\sigma_*^2}{ \sigma_n^2}},
	\\
	\bregmanconj(\m_*; \m_n) 
	&= \half \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \half \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
where all these lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := \half (z - 1 - \log(z)),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n ) 
	:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) 
	= \half \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman}
}
which can also be expressed with $\phi$ as follows:
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{\inv{z}}
	&=  \half \paren{z -1  + \inv{z} -1} 
	= \frac{(z-1)^2}{z} \; .
}
See the grey upper bound in Figure~\ref{fig:phi}.

>>>>

\begin{figure}[ht]
	\centering
\includegraphics[width=.6\textwidth]{gammaMAPrate.pdf}
	\caption{KL divergence~\eqref{eq:suboptimalityKL} for Gaussian variance (\S\ref{ssec:gaussian-variance}) MLE (blue) and MAP (orange) against number of samples $n$. 
		Solid curve  are average over $10^5$ trials.
		Dashed curves are upper bounds~\eqref{eq:MLE_rate} (blue) and~\eqref{eq:MAP_rate} (orange, tight up to a factor 2).
		Shaded areas are 90\% confidence interval.
		The MLE expected KL is infinite for $n=1$ and $n=2$, but for $n\geq3$ it quickly joins the upper bound~\eqref{eq:MLE_rate} and the $1/2n$ asymptote~\eqref{eq:asymptote}.
		MAP's expected KL is always finite, and it has lower variance than MLE, but it is slower to join the asymptote.
	}
	\label{fig:MAP_rate}
\end{figure}

In this section, we prove the results mentioned in \S\ref{ssec:gaussian-variance},
and add some context and experimental observations.
In particular, we show the bound~\eqref{eq:MAP_rate} in \cref{fig:MAP_rate}.
Let us recall the form of the divergence with different parametrizations.
For a Gaussian with known mean $0$ and covariance $\sigma^2$, 
natural and mean parameters $\theta, \m$;
\begin{align}
	\KL( \sigma_*^2 ; \sigma_n^2 ) 
	&= \half \left ( \frac{\sigma_*^2}{ \sigma_n^2} - 1 - \log \frac{\sigma_*^2}{ \sigma_n^2} \right)
	= \phi\paren{\frac{\sigma_*^2}{ \sigma_n^2}},
	\\
	\bregmanconj(\m_*; \m_n) 
	&= \half \left ( \frac{\m_*}{ \m_n} - 1 - \log  \frac{\m_*}{ \m_n} \right) 
	= \phi\paren{\frac{\m_*}{ \m_n}},
	\\
	\bregman( \nat_n; \nat_* ) 
	&=  \half \left ( \frac{ \nat_n}{\nat_*} - 1 - \log  \frac{ \nat_n}{\nat_*} \right)
	= \phi\paren{\frac{ \nat_n}{\nat_*}},
\end{align}
where all these lines are equal, and $\phi$ measures the discrepancy between the ratio $\frac{ \nat_n}{\nat_*} =  \frac{\m_*}{ \m_n}  $ and $1$ via the function
\begin{align}
	\phi(z) := \half (z - 1 - \log(z)),
\end{align}
illustrated in orange in Figure~\ref{fig:phi}.
To upper bound for the MAP, due to the difficulty of finding a closed form for the expectation of the logarithm, we focus on the symmetrized Bregman instead
\alignn{
	\bregmanconj( \mu_*; \mu_n) 
	\leq \cS_{\conj} (\mu_*, \mu_n ) 
	:= \bregmanconj( \mu_*; \mu_n)  + \bregmanconj( \mu_n; \mu_*) 
	= \half \paren{\frac{ \mu_*}{\mu_n} -1 +\frac{ \mu_n}{\mu_*} - 1 },
	\label{eq:symmetrized_bregman}
}
which can also be expressed with $\phi$ as follows:
\aligns{
	\label{eq:log_bound}
	\phi(z) \leq \phi(z) + \phi\paren{\inv{z}}
	&=  \half \paren{z -1  + \inv{z} -1} 
	= \frac{(z-1)^2}{z} \; .
}
See the grey upper bound in Figure~\ref{fig:phi}.
